Boson-Sampling with Single Photons: A Roadmap to Post-Classical Computation

### 2016-10-14 roadmap.tex
added table and figure on post-classical era

To model the above, we must first employ a representation for photons, which accommodates their spatio-temporal structure. Ordinarily, we represent photons as the photon creation operator acting on the vacuum state, \mbox{$\ket{1}=\hat{a}^\dag\ket{0}$}. This is appropriate when dealing with indistinguishable photons. However, photons have rich spatio-temporal structure, and in these degrees of freedom photons are never perfectly indistinguishable. To capture this, we replace photon creation operators with mode operators \cite{bib:RohdeMauerer}, that represent photons in these additional degrees of freedom. In the temporal degree of freedom, for example,
\begin{align}
\hat{A}^\dag_\psi = \int_{-\infty}^\infty \psi(t) \hat{a}^\dag(t) \,dt.
\end{align}
Here $\psi$ is the temporal distribution function (TDF) of the photon, and $\hat{a}^\dag(t)$ is the time-specific creation operator. $\psi$ is normalised, such that,
\begin{align}
\int_{-\infty}^\infty |\psi(t)|^2 \, dt = 1.
\end{align}
The Fourier transform gives us the spectral distribution function. Then, a single-photon state with associated TDF $\psi$, is given by,
\begin{align}
\ket{1_\psi} = \hat{A}^\dag_\psi \ket{0}.
\end{align}
This can be logically generalised into other degrees of freedom, such as the spatial degrees of freedom. For example, to capture both temporal and $x$-$y$ spatial degrees of freedom, we could use the representation,
\begin{align}
\hat{A}^\dag_\psi = \int_{-\infty}^\infty \int_{-\infty}^\infty \int_{-\infty}^\infty \psi(t,x,y) \hat{a}^\dag(t,x,y) \, dt\, dx\, dy.
\end{align}

@@

\begin{figure}[!htb]
%\includegraphics[width=\columnwidth]{moores_law}
\caption{Extrapolation of desktop PC (bottom) and top-end supercomputer (top) performance over time, following an exponential fit, as predicted by Moore's Law.} \label{fig:moores_law}
\end{figure}

From our predictions for the performance of classical computing 5 years into the future, in Fig.~\ref{fig:outperform} we relate these predicted performances to the scattershot boson-sampling protocol, to estimate the number of photons required to match this performance. We further estimate the number of photons required to outperform these benchmarks by a factor of 10, taking us into the regime of post-classical computation.

\begin{figure}[!htb]
%\includegraphics[width=\columnwidth]{outperform}
\caption{Performance of scattershot boson-sampling, showing data-points coinciding with our expectations for desktop PC and supercomputer performance in 5 years, as well as exceeding these anticipated performances by a factor of 10, which we use as a benchmark for the \emph{post-classical era}.} \label{fig:outperform}
\end{figure}

The number of photons required for these different benchmarks are summarised in Tab.~\ref{tab:benchmark_outperform}

\begin{table*}[!htb]
\begin{tabular}{|c|c|c|c|c|}
\hline
Classical computer in 5 years & Classical FLOPS & Single-photon efficiency ($\eta$) & Photon-number ($n$) & Single-photon fidelity ($F$) \\
\hline
\hline
Desktop PC & . & 99\% & . & . \\
 & & 99.9\% & . & . \\
\hline
Desktop PC (10x) & . & 99\% & . & . \\
 & & 99.9\% & . & . \\
\hline
Supercomputer & . & 99\% & . & . \\
 & & 99.9\% & . & . \\
\hline
Supercomputer (10x) & . & 99\% & . & . \\
 & & 99.9\% & . & . \\
\hline
\end{tabular} \caption{Comparison of the classical performance of different classical computers predicted in 5 year's time, their associated performance in FLOPS, respective number of photons required by a boson-sampler to match their performance.} \label{tab:benchmark_outperform}
\end{table*}

>>>>

To model the above, we must first employ a representation for photons, which accommodates their spatio-temporal structure \cite{bib:RohdeRalph05}. Ordinarily, we represent photons as the photon creation operator acting on the vacuum state, \mbox{$\ket{1}=\hat{a}^\dag\ket{0}$}. This is appropriate when dealing with indistinguishable photons. However, photons have rich spatio-temporal structure, and in these degrees of freedom photons are never perfectly indistinguishable. To capture this, we replace photon creation operators with mode operators \cite{bib:RohdeMauererSilberhorn07}, that represent photons in these additional degrees of freedom. In the temporal degree of freedom, for example,
\begin{align}
\hat{A}^\dag_\psi = \int_{-\infty}^\infty \psi(t) \hat{a}^\dag(t) \,dt.
\end{align}
Here $\psi$ is the temporal distribution function (TDF) of the photon, and $\hat{a}^\dag(t)$ is the time-specific creation operator. $\psi$ is normalised, such that,
\begin{align}
\int_{-\infty}^\infty |\psi(t)|^2 \, dt = 1.
\end{align}
The Fourier transform gives us the spectral distribution function. Then, a single-photon state with associated TDF $\psi$, is given by,
\begin{align}
\ket{1_\psi} = \hat{A}^\dag_\psi \ket{0}.
\end{align}
This can be logically generalised into other degrees of freedom, such as the spatial degrees of freedom. For example, to capture both temporal and $x$-$y$ spatial degrees of freedom, we could use the representation,
\begin{align}
\hat{A}^\dag_\psi = \int_{-\infty}^\infty \int_{-\infty}^\infty \int_{-\infty}^\infty \psi(t,x,y) \hat{a}^\dag(t,x,y) \, dt\, dx\, dy.
\end{align}

@@

\begin{figure}[!htb]
\includegraphics[width=\columnwidth]{moores_law}
\caption{Extrapolation of desktop PC (bottom) and top-end supercomputer (top) performance over time, following an exponential fit, as predicted by Moore's Law.} \label{fig:moores_law}
\end{figure}

From our predictions for the performance of classical computing 5 years into the future, we relate these predictions to the scattershot boson-sampling protocol, to estimate the number of photons required to match this performance. We further estimate the number of photons required to outperform these benchmarks by a factor of 10, taking us into the regime of post-classical computation. The number of photons required for these different benchmarks are summarised in Tab.~\ref{tab:benchmark_outperform}.

\begin{table*}[!htb]
\begin{tabular}{|c|c|c|c|c|}
\hline
Classical computer in 5 years & Classical FLOPS & Single-photon efficiency ($\eta$) & Photon-number ($n$) & Single-photon fidelity ($F_\mathrm{in}$) \\
\hline
\hline
Desktop PC & $6.9\times 10^{10}$ & 99\% & 3 & 0.9826 \\
 & & 99.9\% & 3 & 0.9826 \\
\hline
Desktop PC (10x) & $6.9\times 10^{11}$ & 99\% & 5 & 0.9895 \\
 & & 99.9\% & 5 & 0.9895 \\
\hline
Supercomputer & $2.0\times 10^{18}$ & 99\% & 21 & 0.9975 \\
 & & 99.9\% & 21 & 0.9975 \\
\hline
Supercomputer (10x) & $2.0\times 10^{19}$ & 99\% & 24 & 0.9978 \\
 & & 99.9\% & 23 & 0.9977 \\
\hline
\end{tabular} \caption{Comparison of the classical performance of different classical computers predicted in 5 year's time, their associated performance in FLOPS, and respective photon-number required by a scattershot boson-sampler to match their performance. We assume the repetition rate of the boson-sampler to be an experimentally realistic \mbox{$R=76\mathrm{MHz}$}, and impose the requirement that the probability a given sample is taken from a boson-sampling distribution to be \mbox{$F_\mathrm{out}=0.9$}. \comment{Zuen: Please confirm that for your calculatins $\eta$ is the single-photon efficiency, not the net efficiency.}} \label{tab:benchmark_outperform}
\end{table*}
